---
title: "Supplementary Material - VENturing forth: The utility of machine learning
  for the neurometric analysis of von Economo neurons (VENs)"
author: "Bruton et al. 2025"
date: "2025-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Load libraries

```{r eval = FALSE}
library(readr)
library(corregp)
library(MachineShop)
```

# Read data 

Read the DataNeuroMorpho_Human_pyramidal_28-11-24.csv file containing the data 
```{r eval = FALSE}
df <- read_csv(file.choose())
```


# Prepare the data

Convert the column cell_type1 in factor


```{r eval = FALSE}
Human_pyramidal$cell_type1 <- as.factor(Human_pyramidal$cell_type1)
```

# Select variables

Select morphological variables that will work as predictors. Dataframe number columns from 2 to 24.

```{r eval = FALSE}
Variables <- names(Human_pyramidal)[2:34]
```
# Bootstraping

Estimate variable importance from six different machine learning models using 5000 sub-samples of 10 variables/predictors. 

```{r eval = FALSE}
Results <- NULL
for (i in c(1:5000)) {

  Vars <- sample(Variables, 10)
  formula <-  as.formula(paste('cell_type1 ~ ', Vars[1], "*",Vars[2], "*", Vars[3], "*", Vars[4], "*", Vars[5], "*", Vars[6], "*", Vars[7], "*", Vars[8], "*", Vars[9], "*", Vars[10] , sep=""))
  
  
  model_1 <- fit(formula, data = Human_pyramidal, model = EarthModel)
  vars_earth <- varimp(model_1)
  
  # this will take some time:
  model_2 <- fit(formula, data = Human_pyramidal, model = BARTModel)
  vars_bart <- varimp(model_2)
  
  model_3 <- fit(formula, data = Human_pyramidal, model = C50Model)
  vars_C50 <- varimp(model_3)
  
  model_4 <- fit(formula, data = Human_pyramidal, model = SVMModel)
  vars_SVM <- varimp(model_4)
  
  model_5 <- fit(formula, data = Human_pyramidal, model = XGBModel)
  vars_XBG <- varimp(model_5)
  
  model_6 <- fit(formula, data = Human_pyramidal, model = RandomForestModel)
  vars_RF <- varimp(model_6)
  
  # extracting results form each model
  Temp1 <- list(names=vars_earth@row.names, values=vars_earth@.Data[1] )
  Temp1 <- as.data.frame(Temp1)
  names(Temp1) <- c("Variables", "Values")
  Temp1["Order"] <- c(1:10)
  Temp1["Model"] <- "EarthModel"
  Temp1["Repetition"] <- i
  
  Temp2 <- list(names=vars_C50@row.names, values=vars_C50@.Data[1] )
  Temp2 <- as.data.frame(Temp2)
  names(Temp2) <- c("Variables", "Values")
  Temp2["Order"] <- c(1:10)
  Temp2["Model"] <- "BARTModel"
  Temp2["Repetition"] <- i
  
  Temp3 <- list(names=vars_bart@row.names, values=vars_bart@.Data[1] )
  Temp3 <- as.data.frame(Temp3)
  names(Temp3) <- c("Variables", "Values")
  Temp3["Order"] <- c(1:10)
  Temp3["Model"] <- "C50Model"
  Temp3["Repetition"] <- i
  
  Temp4 <- list(names=vars_SVM@row.names, values=vars_SVM@.Data[1] )
  Temp4 <- as.data.frame(Temp4)
  names(Temp4) <- c("Variables", "Values")
  Temp4["Order"] <- c(1:10)
  Temp4["Model"] <- "SVMModel"
  Temp4["Repetition"] <- i
  
  Temp5 <- list(names=vars_XBG@row.names, values=vars_XBG@.Data[1] )
  Temp5 <- as.data.frame(Temp5)
  names(Temp5) <- c("Variables", "Values")
  Temp5["Order"] <- c(1:10)
  Temp5["Model"] <- "XGBModel"
  Temp5["Repetition"] <- i
  
  Temp6 <- list(names=vars_RF@row.names, values=vars_RF@.Data[1] )
  Temp6 <- as.data.frame(Temp6)
  names(Temp6) <- c("Variables", "Values")
  Temp6["Order"] <- c(1:10)
  Temp6["Model"] <- "RandomForestModel"
  Temp6["Repetition"] <- i
  
  Results <- rbind(Results, Temp1, Temp2, Temp3, Temp4, Temp5, Temp6)
  
}
```

